Packet Optical Transport Network Slicing with Hard and Soft Isolation

### 2021-06-06 template.tex
Updates from Overleaf

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
\usepackage{cleveref}
\usepackage{overpic}
%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

@@

% Contact information of the corresponding author
\corres{Correspondence: samier.barguil@estudiante.uam.es)}

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{(Victor) We validate the deployment of isolated transport network slices in IP over DWDM network. To this end, an isolated transport network slice is deployed using multi-layer isolation mechanisms based on OpenConfig and ONF Transport API.
\\A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusion: indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}

@@

\section{Introduction - Luis}

@@

%This paper presents an end-to-end architecture to provide transport network slices deployed over multi-layer IP over DWDM networks. Several degrees of isolation (from hard to soft) might be required and implemented in the requested transport network slice. This is the first paper to explore transport network slice isolation using an IP over DWDM network. In order to validate this proposed architecture, we present a proof-of-concept in Telefonica and CTTC Laboratories.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Network Slicing Concepts}

In the context of transport networks, the NGMN defines two main concepts\cite{alliance2016description}: 
\begin{itemize}
    \item The \textit{Service Instance} (SI) is an end-user service or a business service realized in a Network Slice.
    \item The \textit{Network Slice Instance} (NSI) as the complete, instantiated logical network that meets specific characteristics required by a Service instance.
\end{itemize}
Hence, network slicing is sharing network infrastructure across different Service Instance(s) to meet network-specific requirements. Depending on the communication layer where a network operator implements network slicing, resource management, the network characteristics or the toolbox used to implement the network slice can change. Some examples, of the network characteristics requested by a Service Instance can be ultra-low-latency or ultra-reliability, etc \cite{alliance2016description,}.

The network slicing concept provides a framework for broad applicability across various industries. The majority of these scenarios envisioned to suit emerging and diverse business models based on the Network As A Service (NASS) approach, creating opportunities for intelligent services and a new business ecosystem \cite{kuklinski2019business,zhang2017network}., One of the main enablers to drive the network slicing is the 5G networks realization.

5G supports a new generation of applications with diverse and stringent requirements in terms of capacity, latency, level of mobility, number of users, and user density. 5G actually defines three types of applications based on how their data needs to be treated:

\begin{itemize}
\item Ultra Reliable Low Latency Communications (URLLC)
\begin{itemize}
\item Requires support for 1 ms latencies, 0.001% packet loss, user mobility up to 100km/h
\item E.g., autonomous driving
\end{itemize}
\item Massive Machine Type Communications (mMTC)
\begin{itemize}
\item Requires support for 1 million devices per square kilometer, tens of bps bandwidth, and latency minimization for battery life optimization
\item E.g., Massive IoT
\end{itemize}
\item Enhanced Mobile Broadband (eMBB)
\begin{itemize}
\item Requires Gbps bandwidth, real time or not
\item  E.g., immersive UIs based on Augmented Reality / Virtual Reality
\end{itemize}
\end{itemize}

\subsection{Soft Network Slicing}

Network slicing at the MPLS level can be implemented using Virtual Routing and Forwarding (VRF), that enables multiple routing environments over a shared MPLS transport network, and Virtual switching Interfaces (VSI), that enables multiple switching environments over the same shared infrastructure. Each physical router is able to host multiple VRFs and multiple VSIs (along with their attached logical interfaces), effectively slicing it into multiple routing and switching environments that can be assigned to different tenants.

This slicing of a shared MPLS transport network into multiple routing/switching environments assigned to different tenants, with MPLS paths in each environment subject to potentially different QoS treatment, is known as soft network slicing.

\subsection{Hard Network Slicing}

VRFs and VSIs cannot be managed by their corresponding tenants directly because they are part of the same administrative domain represented by the physical router, and they instead need to be managed by the operator that owns the physical routing infrastructure.  
Disaggregated routers, with data plane running in the physical device, and with the control plane running outside the device in a remote cloud or remote server, are able to disassociate the 1:1 relationship between physical device and router, thus enabling the support of multiple virtual routers over a single physical network device.

Each of these virtual routers is a router in its full sense, being able to host multiple VRFs and VSIs, or to be managed independently of the other virtual routers running in the same device. Virtual routers are thus administratively separated from each other, which means separate virtual routers can be assigned to different tenants, and each tenant can manage the virtual routers directly, without the need from the operator that owns the physical network to intervene.

Network slicing at the IP transport level can then be accomplished with grouping multiple virtual routers running over a shared physical network infrastructure into a common virtual infrastructure under its own separate administrative domain. The set of virtual routers under the same administrative domain is then known as hard network slicing, and mechanisms such as Role Based Access Control (RBAC) can be used to assign different hard slices to different tenants.

Finally, technologies such as FlexE can also be used to enable separate physical connectivity for different hard network slices sharing the same devices and interconnecting links.

@@

A 5G cellular network has three main components: the radio access network (RAN), the mobile core (MC) and the backhaul network that interconnects the RAN with the mobile core.

\begin{itemize}
\item  The Radio Access Network (RAN)  is specified within 5G, and it covers everything related to the air interface between the user element and the base station.

\item  The Mobile Core (MC) is also specified within 5G, and its main role is to act as a gateway for user traffic to and from the greater internet (user plane), to manage user mobility, authentication, registration, etc (control plane), and to establish per-user tunnels between user plane and base stations for each different traffic type.

\item The Backhaul Network is the network that interconnects the RAN with the MC. It is not part of the 5G specification, so it is up to each network operator to decide how to implement it. It requires functionalities such as QoS, timing synchronization, MPLS, and segment routing.
\end{itemize}

The network slicing described so far applies to the mobile backhaul network, but not RAN or the MC. Instead, 3GPP defined the Network slicing in 5G, specifying how network operators can use distributed software components to implement virtual schedulers in RAN and service meshes in Mobile Core.

5G enables network operators to ensure the same network can fulfill the heterogeneous requirements of these diverse types of applications by determining how network resources are assigned to application traffic. This is defined as network slicing. 5G actually specifies a standard set of network slices, denoted Standardized Slice Type (SST), in order to determine how resources need to be assigned at the RAN and at the MC level to fulfill the requirements of different types of applications. For example, SST1 applies to eMBB, SST2 applies to URLLC, and SST3 to mMTC.

Network slicing capabilities need to be available across all components of the 5G cellular network (RAN, MC, Backhaul network) in order to ensure this differentiated treatment end to end, for example from the moment user traffic enters the RAN to the moment it exits the MC user plane towards the Internet. 5G specifies how network slicing is implemented in RAN and MC, but it does not specify how the backhaul network achieves it, so it is up to each network operator to determine how to accomplish it.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Instantiating of Network Slices in SDN transport networks - Oscar}

\subsection{Hierarchical Network Controller}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Proposed architecture - Victor}

%Figure \ref{fig:scheme}.a shows a network scheme of the approached concept. It consists of three main domains: SDN domain, IP domain, and optical domain.

%The Network Slice Controller (NSC) realizes a transport network slice in the underlying transport infrastructure, maintains and monitors the  state of its resources.  The NSC will delegate on SDN Domain controllers \cite{ifusion} to configure the network resources. The NSC receives a transport network slice request from the Operation Support System and Business Support System (OSS/BSS). The request is modeled uing the YANG data model defined in \cite{transportslicemodel} by means of the RESTCONF protocol \cite{restconf}. Internal workflow for transport network slice life-cycle management is prepared on top of L2/L3 service management (SM) workflows. In order to interact with underlying IP and Optical Domain controllers via a RESTCONF client.

\begin{figure}[tb]%[htdp]
\centering
	\centering
		\includegraphics[width=\linewidth]{arch.png}
\vspace{-7pt} \caption{a) Overall proposed architecture; and b) Sequence diagram for multiple network slice deployment with isolation characteristics.}
\label{fig:scheme}
\vspace{-17pt}
\end{figure}

%The OSS/BSS may request that the deployed network slice is isolated from any other network slices or different services delivered to the same customers. Naturally, other network slices or services must not negatively impact the requested transport network slice's delivery. There are several possibilities to provide this isolation, which can be provided at several degrees, such as dedicated allocation of resources for a specific slice or sharing some network resources. Figure \ref{fig:results}.a shows the multiple isolation options that range from a hard slice to a soft slice: a) no-isolation, meaning that slices are not separated; b) physical-isolation, where slices are completely physically separated, for example, in different locations; c) logical-isolation, where slices are logically separated, only a certain degree of isolation is performed through QoS mechanisms; d) process-isolation, where slices include process and threads isolation; e) physical-network-isolation, where slices contain physically separated links; f) virtual-resource-isolation, where slices have dedicated virtual resources; g) network-functions-isolation, where Network Function (NF) are dedicated to a single network slice; h) service-isolation, where virtual resources and NFs are shared.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

%The experimental testbed has been set-up at Telefónica (NSC deployment and Volta Elastic Virtual Routing Engine stacked in 7316 Edgecore hardware) and CTTC laboratories (including a T-API SDN controller over a flexi-grid 4-nodes DWDM network), as depicted in Figure \ref{fig:scheme}.a. In order to properly set-up transport, we have introduced the transport network slice YANG data model in the RESTCONF server of NSC. An example of requested transport network slice is depicted in Figure \ref{fig:results}.c. It can be observed that several levels of isolation can be requested per node, link and slice.

%In order to validate the depicted sequence diagram in Figure \ref{fig:scheme}.b, we provide the captured Wireshark traces among the complete system to deploy a complete hard slice request (the figure provides the selected significant traces). In this figure the different involved protocols can be observed. Firstly it shows the request for the transport network slice. It is quickly stored and answered, as it is processed after response (later an status update might be requested).

>>>>

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
\usepackage{cleveref}
\usepackage{enumitem}
%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

@@

% Contact information of the corresponding author
\corres{Correspondence: samier.barguil@estudiante.uam.es}

@@

% Abstract (Do not insert blank lines, i.e. \\)

\abstract{Network Operators has been dealing with the necessity of a dynamic network resources allocation to provide a new generation of customer-tailored applications. In that sense, Telecom providers have to migrate their BSS/OSS systems and network infrastructure to more modern solutions to introduce end-to-end automation and support the new use cases derived from the 5G adoption and transport network slices. In general, there is a joint agreement on making this transition to an architecture defined by programmable interfaces and standard protocols. Hence, this paper uses the iFusion architecture to control and program the network infrastructure. The work presents an experimental validation of the network slicing instantiation in an IP/Optical environment using a set of standard protocols and interfaces. The work provides results of the creation, modification and deletion of the network slices. Furthermore, it demonstrates the usage of standard communication protocols (Netconf and Restconf) in combination with standard YANG data models.}
%A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusion: indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{}

@@

\section{Introduction}

Network slicing is positioned as the next paradigm for service delivery in telecom operator networks. the slicing concept departs from the idea of allocating network resources to customer services on demand, leverages on the paradigms of Software Defined Networking (SDN) and Network Function Virtualization (NFV) \cite{ordonez2017network}, being those resources dedicated to them during service lifetime. Those assigned resources (including compute, storage and transport ones) can be either physical or virtual, tailored to the customer need expressed through the slice request. Since distinct customers could have different requirements to be satisfied, different slice types can be considered from the point of view of management and control, as defined in \cite{contreras2017network}.

The way of slicing can be seen as an alternative form of consuming network capabilities, permitting to address the particular needs of new industries and markets \cite{ordonez2017network} as was not possible before. Specific service characteristics like deterministic extreme low latency or guaranteed high bandwidth, different degrees of isolation (with respect other customers’ services), scalability adapted to the actual demand, or  resource management and control can be now enabled, allowing more sophisticated services to emerge and co-exist in a commonly shared infrastructure.

In order to enable this flexible consumption of network capabilities, it is necessary to develop new forms of network control to make possible a dynamic partitioning and assignment of resources end-to-end. Therefore, several initiatives have been discussed across the industry, including the IP network resources management (L2 \cite{barguil2021l2nm} and L3VPNs \cite{Aguado.2021l3nm}), the optical layer management (Transport API \cite{vilalta2018experimental}) or the network devices configuration (Openconfig \cite{shaikhopenconfig}).

@@

This paper extends the work presented in the lastest OFC conference \cite{OFC2021alcala}. In addition, this work adds a set of tests; Including creating two isolated network slices, prefixes addition and modification to each slice and, slices deletion. A last set of tests included the service continuity evaluation after a physical device reset. For the whole set of experiments, a hierarchical control architecture over multi-layer IP over DWDM networks was used.

%Several degrees of isolation might be required and implemented in the requested transport network slice.

This paper is organized as follows. The authors describe those concepts in \cref{sec:slicing}. \Cref{sec:arch} described the whole control architecture used for the network slice instantiation in a service provider environment. \Cref{sec:instantiation} describes the proposed architecture's implementation choices. The authors selected one of those choices and made a proof-of-concept in Telefonica and CTTC Laboratories, \cref{sec:results} includes the results.

The list of abbreviations and its definition is defined in \cref{tab:abbreviations}.

\begin{table}[htb!]
\caption{List of abbreviations used across the document}
\label{tab:abbreviations}
\begin{tabular}{|l|l|}
\hline
Abbreviation & \multicolumn{1}{c|}{Definition}              \\ \hline
3GPP         & 3rd Generation Partnership Project           \\ \hline
API          & Application programming interface            \\ \hline
IETF         & Internet Engineering Task Force              \\ \hline
eMBB.        & Enhanced Mobile Broadband                    \\ \hline
L2SM         & L2VPN Service Model                          \\ \hline
L2NM         & L2VPN Network Model                          \\ \hline
L3SM         & L3VPN Service Model                          \\ \hline
L3NM         & L3VPN Network Model                          \\ \hline
MC           & Mobile Core                                  \\ \hline
mMTC         & Massive Machine Type Communication           \\ \hline
MPLS         & Multiprotocol Label Switching                \\ \hline
NASS         & Network As A Service                         \\ \hline
NGMN         & Next Generation Mobile Networks              \\ \hline
NSI          & Network Slice Instance                       \\ \hline
ONF          & Open Networking Foundation                   \\ \hline
OSS          & Operation Support Systems                    \\ \hline
QoS          & Quality of service                           \\ \hline
RAN          & Radio Access Network                         \\ \hline
SDN          & Software Defined Network                     \\ \hline
SI           & Service Instance                             \\ \hline
URLLC        & Ultra Reliable Low Latency Communications    \\ \hline
VSI          & Virtual switching Interfaces                 \\ \hline
VRF          & Virtual Routing and Forwarding.              \\ \hline
YANG         & Yet Another Next Generation                  \\ \hline
\end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Network Slicing Concepts}
\label{sec:slicing}

Network slicing refers to the partitioning of one physical network into multiple virtual networks, each network slice is architected and optimized for a specific application/service.  In this context, the Next Generation Mobile Networks (NGMN) defines two main concepts \cite{alliance2016description}: 
\begin{itemize}
    \item The \textit{Service Instance} (SI) is an end-user service or a business service realized in a Network Slice.
    \item The \textit{Network Slice Instance} (NSI) as the complete, instantiated logical network that meets specific characteristics required by a Service instance.
\end{itemize}

Hence, network slicing is sharing network infrastructure across different Service Instance(s) to meet network-specific requirements. Depending on the communication layer where a network operator implements network slicing, resource management, the network characteristics or the toolbox used to implement the network slice can change. Some examples, of the network characteristics requested by a Service Instance can be ultra-low-latency or ultra-reliability, etc \cite{alliance2016description}.

The network slicing concept provides a framework for broad applicability across various industries. The majority of these scenarios envisioned to suit emerging and diverse business models based on the Network As A Service (NASS) approach, creating opportunities for intelligent services and a new business ecosystem \cite{kuklinski2019business,zhang2017network,khan2020network}., One of the main enablers to drive the network slicing is the Fifth-generation (5G) networks realization. Due to the necessity to integrate multiple services with various performance requirements — such as high throughput, low latency, high reliability, high mobility, and high security — into a single physical network infrastructure, and provide each service with
a customized logical network. being the Network Slicing the key technology to achieve the aforementioned goals.

Due to the 5G expectation of supporting of a new generation of customer-tailored applications with diverse and stringent requirements (e.g., in terms of capacity, latency, level of mobility, number of users, and user density), 5G sets the stage for innovation and transformation in consumer services and vertical industries, such as real-time services, location-based services, and mission-critical services, as the ones described in the \cref{TAB1} \cite{foukas2017network}.

\begin{table}[htb!]
\caption{TAB1:}
\label{TAB1}
\centering
\begin{tabular}{|l|l|l|}
\hline

\textbf{Applications} & \textbf{Definition}& \textbf{Example} \\ \hline \hline

\begin{tabular}[c]{@{}l@{}}Ultra Reliable Low \\ Latency Communications \\ (uRLLC)\end{tabular} & \begin{tabular}[c]{@{}l@{}}Requires support for 1 ms \\ latencies, 0.001\% packet loss,\\ user mobility up to \\100km/h\end{tabular}  & \begin{tabular}[c]{@{}l@{}} Autonomous driving or \\ Industrial automation \end{tabular} \\ \hline

\begin{tabular}[c]{@{}l@{}}Massive Machine \\ Type Communications \\ (mMTC)\end{tabular}   & \begin{tabular}[c]{@{}l@{}}Requires support for 1 million \\ devices per square kilometer, \\ tens of bps bandwidth, \\ and latency minimization \\ for battery life optimization\end{tabular} & Massive IoT \\ \hline

\begin{tabular}[c]{@{}l@{}}Enhanced Mobile \\ Broadband (eMBB)\end{tabular} & \begin{tabular}[c]{@{}l@{}}Requires Gbps bandwidth, \\ real time or not \end{tabular}& \begin{tabular}[c]{@{}l@{}}Immersive UIs based \\ Augmented Reality \\ Virtual Reality.\\ Streaming of \\ High Quality Video\end{tabular} \\ \hline
\end{tabular}
\end{table}

The services described in \cref{TAB1} demonstrate highly diversified traffic characteristics and differentiated quality-of-service (QoS) requirements. For example, the mMTC is based on its application in machine-to-machine communication \cite{huang20205g}. The actual realization of those services in telecom networks implies the migration of the current heterogeneous IP/MPLS + DWDM transmission networks to more modern and 5G-ready designs, using the Network programmability, Software-Defined Networking (SDN) and Machine Learning (ML) as their main pillars.

Implementing the network slices has two main alternatives: Hard and Soft. The way the network resources are shared between the services in each alternative is described in the following sections.

\subsection{Soft Network Slicing}
\label{sec:soft}

Soft slicing corresponds to a lower level of isolation between the services a network is transporting. Soft slicing implies sharing the physical infrastructure but creates logical segmentations between the customers. According to this definition, soft slicing is not a new concept from the IP/MPLS networks perspective \cite{rosen2006rfc,l3sm2017Litkowski,Aguado.2021l3nm}.

Traditional L3VPNs are samples of soft-slicing implementations in an MPLS network, because a VPN can be thought of as a series of tunnels connecting customer sites, each site can potentially have different QoS treatment, and all traffic to and from each site is internal to the customer. In the VPN service, the provider ensures that each customer's traffic is logically discriminated over shared physical infrastructure based on routing policies configuration across the network.

In that sense, Network slicing at the MPLS level can be implemented using: 
\begin{itemize}
    \item \textbf{Virtual Routing and Forwarding} (VRF), that enables multiple routing environments over a shared MPLS transport network.
    \item \textbf{Virtual Interfaces} (VSI), that enables multiple switching environments over the same shared infrastructure. 
\end{itemize}

Each physical router is able to host multiple VRF(s) and multiple VSI(s) (along with their attached logical interfaces), effectively slicing it into multiple routing and switching environments that can be assigned to different tenants/customers/services.

%In the optical network the concept is also well known and the soft-slicing can be mapped with the and virtual containers and optical data unit (ODU) running on top of a L0/L1 shared underlay.

\subsection{Hard Network Slicing}
\label{sec:hard}

Despite the apparent advantages that an overlay MPLS tunnels could provide, there are some disadvantages. Overlay tunnels built by data encapsulation have neither visibility nor control of the underlying physical network. With more and more tunnels deployed on shared physical infrastructure, network congestion necessarily becomes an attention point.  Thus, VRFs, VSIs or ODus cannot be managed by their corresponding tenants directly because they are part of the same administrative domain represented by the physical device, and they need to be operated by the network administrator. Due to the massive number of tunnels and services in a complex network, the QoS and Traffic Engineering policies management becomes a crucial task to guarantee the right SLAs to its customers \cite{huang20205g,marotta2020reliable}.

The solution to this limitation and the way to move the networking industry to more automated scenarios is Hard slicing. Hard-slicing refers to the provision of dedicated resources to a specific Network Slice Instance. For example, Data-plane resources are provided by allocating time-domain multiplexed resources such as a Flex Ethernet channel or as a service such as an MPLS hard-pipe, route diversity (disjoint paths), wavelength selection, among others.

Disaggregated routers is an example of hard-slicing. In the disaggregated case, the data plane runs in the physical device, and the control plane runs outside the device in a remote cloud or server. In that case, the 1:1 relationship between physical device and routing logic is disassociated, enabling the support of multiple virtual routers over a single physical network device. Each of these virtual routers is a router in its full sense, being able to host multiple VRFs and VSIs, or to be managed independently of the other virtual routers running in the same device. Virtual routers are thus administratively separated from each other, which means separate virtual routers can be assigned to different tenants, and each tenant can manage the virtual routers directly, without the need from the operator that owns the physical network to intervene intervene (e.g., following an approach similar to the one described in \cite{contreras2021architecture})

Network slicing at the IP transport level can then be accomplished by grouping multiple virtual routers running over a shared physical network infrastructure into a common virtual infrastructure under its separated administrative domain. Virtual routers under the same administrative domain are then known as hard network slice, so different hard slices can be assigned to different tenants.

@@

5G must enable network operators to ensure the same network can fulfil the heterogeneous demands of diverse types of applications. To efficiently achieve those requirements and determining how network resources are assigned, a service provider must integrate technologies like Software-Defined Networking (SDN), Network Functions Virtualization (NFV), and Machine Learning (ML)  with a hierarchical transport architecture.

According to the 5G definition the architecture has three main components \cite{etsi2018123,redana20195g}:

\begin{itemize}
\item  \textbf{Radio Access Network (RAN)}: It covers everything related to the air interface between the user element and the base station. The RAN interfaces and interconnections is specified within 3G PPP Architecture Working Group.

\item  \textbf{Mobile Core (MC)}: Its central role is to act as a gateway for user traffic to and from the internet. The mobile core is composed of a set of network functions (NF) responsible for managing user mobility, access authentication, access authorization, location service management, registration, and establish per-user tunnels between base stations for each different traffic type.

\item \textbf{Backhaul Network} is the network that interconnects the RAN with the MC. It is not part of the 5G specification, so it is up to each network operator to decide how to implement it. It requires functionalities such as QoS management,  synchronization and a stack of protocols like IP/MPLS or segment routing.
\end{itemize}

Network slicing capabilities need to be available across all components of the 5G cellular network (RAN, MC, Backhaul network) in order to ensure a differentiated treatment of the packets in the network. Thus, the 5G working groups have specified a standard set of network slices, denoted Standardized Slice Type (SST), to determine how resources should be assigned at the RAN and MC \cite{ferrus2018automation}.   
%\begin{itemize}
%    \item \textbf{SST}1 applies to eMBB 
%    \item \textbf{SST}2 applies to URLLC
%    \item \textbf{SST}3 to mMTC.
%\end{itemize}

%for example from the moment user traffic enters the RAN to the moment it exits the MC user plane towards the Internet. 5G specifies how network slicing is implemented in RAN and MC, but it does not specify how the backhaul network achieves it, so it is up to each network operator to determine how to accomplish it.

@@

%The network slicing described so far applies to the mobile backhaul network, but not RAN or the MC. Instead, 3GPP defined the Network slicing in 5G, specifying how network operators can use distributed software components to implement virtual schedulers in RAN and service meshes in Mobile Core.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Proposed architecture}
\label{sec:arch}

The i\uppercase{FUSION} architecture is an architecture defined by Telefonica to strengthen the network automation and programmability in a service provider environment, as depicted in \cref{fig:architecture} \cite{contreras2019ifusion}. i\uppercase{FUSION} is a two-layer control architecture, with specific domain controllers per technological domain (IP/MPLS, microwave and optical) in the bottom and a Software-Defined Transport Network controller (SDTN controller) to handle the multi-layer and multi-domain transport network resources.  The domain controllers directly communicate with the network elements and the SDTN controller with the OSS/BSS systems. More than the functional block definition, i\uppercase{FUSION} includes the usage of: (1) Standard interfaces based on \uppercase{RESTconf/YANG} \cite{bierman2017restconf} for the communication between control components and \uppercase{NETCONF/YANG} \cite{enns2011network} to configure the network elements; (2) YANG data models based on latest releases in the standards-development organizations (SDOs): \uppercase{IETF}, \uppercase{ONF} and OpenConfig.

\begin{figure}[tb]%[htdp]
\centering
	\centering
		\includegraphics[width=\linewidth]{Figs/diagram-1.png}
\caption{Overall proposed architecture}
\label{fig:architecture}
\end{figure}

\Cref{fig:architecture} shows the network scheme of the i\uppercase{FUSION} architecture in terms of components and relationships among them. The following sections define each of the structural pillars in the architecture, including its role in network slicing.

%\begin{figure}[tb]%[htdp]
%\centering
%	\centering
%		\includegraphics[width=\linewidth]{Figs/arch.png}
%\caption{a) Overall proposed architecture; and b) Sequence diagram for multiple network slice deployment with isolation characteristics.}
%\label{fig:scheme}
%\end{figure}

@@

\subsection{Software Defined Transport Network Controller} 
\label{sec:ifusionarch}

The Software-Defined Transport Network Controller is a functional block with the following purposes \cite{contreras2019ifusion}; It is The main entry point from the OSS/BSS systems to the network. It is in charge of coordinating/providing services through several domains and layers. It has the multi-layer/multi-domain topological view of the network. The SDTN controller can split requirements based on the technological requirements. During this process, the SDTN controller can add/assign logical resources to be used by the network at the service implementation. The SDTN controller have two RESTCONF interfaces, one to process the OSS/BSS systems' requirements and one to send the specific requests to the domain controllers. TeraFlow project \cite{vilalta21eucnc} is proposing the development of a cloud-native SDN controller that will serve as SDTN.

The SDTN controller can have incorporated the Network Slice Controller as a working piece of its implementation.

The Network Slice Controller (NSC) effectuates a transport network slice in the underlying transport infrastructure, manages and control the state of resources and topologies associated with it. The NSC receives a transport network slice request from the Operation Support System and Business Support System (OSS/BSS). The NSC runs an internal workflow for transport network slice life-cycle management and interacts with underlying IP and Optical Domain controllers via a RESTCONF client. As described in the following sections, depending on the NSC location in the architecture, the NSC will delegate to SDN Domain controllers to configure the network (\cref{sec:instantiation}.a and \cref{sec:instantiation}.b) or make it directly through a NETCONF SBI (\cref{sec:instantiation}.c).

The network slice controller is the key building block for control and management of network slice. It provides the creation/modification/deletion, monitoring an optimization of network slices in a multi-domain, a multi-technology and multi-vendor environment. It has two main functionalities defined in \cite{contreras2020slice,slicing2021framework}:
\begin{itemize}
    \item \textit{Map}: The NSC must Map the Network Slice requests to the underlying technology-specific infrastructure. Accordingly, It maintains a record of the mapping from user requests to slice instantiations, as needed to allow for subsequent control functions like modification or deletion.
    \item  \textit{Realize}: The NSC should realize the network slice request using its SBI interface against the domain controllers in either physical or logical connectivity through VPNs or various tunnelling technologies such as Segment Routing, MPLS, etc. 
\end{itemize}

\subsection{Network Domain Controller}

The network domain controller (SDN Controller) is in charge of network elements (network domain). It has standard southbound interfaces to communicate with the network elements. The Domain controller SBI relies on using the Network configuration protocol (NETCONF) to interact with the underlying technology's network elements. The SDN controller also has a northbound interface to communicate with the SDTN controller or the OSS/BSS Systems using RESTCONF.

\subsection{Yang Models for Network Controllers}

As described until now, the three control elements: SDTN controller, Network Slicing Controller and Network Domain controllers, has standard SBI and NBI interfaces to communicate between them and against the network or the OSS/BSS systems. The standard interfaces are composed of selecting a protocol to transfer the data and YANG data models to define how the message is formed. For that sense, the YANG modelling activities have acquired significant relevance across the standardization entities. To such an extent that by 2019, the Number of correctly extracted YANG models from IETF drafts was 283, in the Broadband forum was 214, and in Openconfig was 137 \cite{benoit2020yang}; similarly, other organizations like the MEF, 3GPP or ONF produced YANG data models to describe technologies, protocols or connectivity services as well. Thus, navigating through the massive set of YANGs available and selecting the suitable pack of data models to define each functional block's interfaces becomes an essential task from the architectural definitions point of view.

To request and after instantiate the network slices a set of these data models are described as the experimental base of this work are described in the \cref{TAB2:models}.

\begin{table}[htb]
\caption{Set of these data models are described as the experimental base of this work}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Models}  & 
\textbf{Description}  & 
\textbf{Example} \\ \hline \hline 
LxVPN  & 
\begin{tabular}[c]{@{}l@{}}These models describe a VPN service \\ from the customer or network \\ operator point of view.\\\end{tabular} &
\begin{tabular}[c]{@{}l@{}}L3SM: \cite{l3sm2017Litkowski} \\ L2SM: \cite{wen2018yang}\\ L3NM:\cite{Aguado.2021l3nm} \\ L2SM: \cite{barguil2021l2nm}\end{tabular} \\ \hline

Traffic Engineering & 
\begin{tabular}[c]{@{}l@{}}These models allow to manipulate \\ Traffic Engineering tunnels \\ within the network segment. \\ Technology-specific extensions \\ allow to work with a desired technology \\ (e.g. MPLS RSVP-TE tunnels, \\ Segment Routing paths,\\OTN tunnels, etc.)\\\end{tabular} & 
\begin{tabular}[c]{@{}l@{}}TE:\cite{te2021Saad}\\ TE Topology:\cite{ietf2021tetopology,ietf2021tel3topology}\end{tabular} \\ \hline

\begin{tabular}[c]{@{}l@{}}TE Service Mapping \\ extensions\end{tabular} & \begin{tabular}[c]{@{}l@{}}These extensions allow to specify \\ for LxVPN the details of an underlay \\ based on Traffic Engineering\\\end{tabular}& 
\begin{tabular}[c]{@{}l@{}}Service \\Mapping: \cite{teas2021servicemapping} \end{tabular} \\\hline

\begin{tabular}[c]{@{}l@{}}ACLs and \\ Routing policies\end{tabular}    & \begin{tabular}[c]{@{}l@{}}Even though ACLs and routing policies\\ are device models,\\It's exposure in the NBI of a domain\\controller allows to provide\\an additional granularity\\that the network domain\\controller is not able to infer on its own.\end{tabular}                 & \begin{tabular}[c]{@{}l@{}}ACL: \cite{ietf2019acl}\\ Routing\\ Policy:\cite{ietf2021routingpolicy,ietf2021bgppolicy}\end{tabular}\\ \hline

OTN  & \begin{tabular}[c]{@{}l@{}}As a part of the transport network,\\
   OTN can provide hard pipes \\with guaranteed data isolation and\\
   deterministic low latency, \\which are highly demanded in the Service\\
   Level Agreement (SLA).\\\end{tabular} & \begin{tabular}[c]{@{}l@{}}OTN Slice: \cite{ietf2021otnslicing}\end{tabular}\\ \hline

Slicing  & \begin{tabular}[c]{@{}l@{}}Set of data models available to map \\ and realize the network Slices.\\ \end{tabular} & \begin{tabular}[c]{@{}l@{}}Network Slice\\ NBI:\cite{slicing2021framework,contreras2020slice}\\ \end{tabular} \\ \hline
\end{tabular}
\label{TAB2:models}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Instantiating of Network Slices in SDN transport networks}
\label{sec:instantiation}

As described before in the \cref{sec:ifusionarch}, the OSS/BSS systems may request the deployment of a new network slices with certain transport characteristics. Each network slice must be isolated from any other network slices or different services delivered to particular customers and naturally, other network slices or services must not negatively impact the requested transport network slice's delivery.

To provide this isolation and instantiate the slice in the network there are certain implementation options ranging from softer to hardest grades of isolation, as follow:
